---
title: "STA 380 Part 2 Exercises - Juwon Lee"
output:
  html_document:
    df_print: paged
---

# Probability Practice

## Part A.

Given information:

* p(random) = 0.3
* p(truthful) = 0.7
* p(yes) = 0.65
* p(no) = 0.35

* p(yes|random) = 0.3 * 0.5 = 0.15
* p(no|random) = 0.3 * 0.5 = 0.15

We want to figure out p(yes|truthful)

* p(yes|truthful) = p(yes and truthful)/p(truthful)
* p(yes) - p(yes and random) = p(yes and truthful)
* 0.65 - 0.15 = 0.5

Plug in:

* p(yes|truthful) = 0.5/0.7

```{r}
0.5/0.7
```

## Part B.

Given information:

* p(test positive|has disease)  = 0.993
* p(test negative | doesn't have disease) = 0.9999
* p(has disease) = 0.000025

Therefore:

* p(has disease and tests positive) = 0.000025 * 0.993
* p(has disease and tests negative) = 0.000025 * 0.007
* p(does not have disease and tests positive) = 0.999975 * 0.0001
* p(does not have disease and tests negative) = 0.999975 * 0.9999
```{r}
0.000025 * 0.993
0.000025 * 0.007
0.999975 * 0.0001
0.999975 * 0.9999
```
We want to figure out p(has disease | tests positive)

* p(has disease | tests positive) = p(has disease and tests positive)/p(tests positive)

* p(tests positive) = p(does not have disease and tests positive) + p(has disease and tests positive)

  = 0.000024825 + 0.000099975 = 0.0001248
  
Therefore:

* p(has disease | tests positive) = p(has disease and tests positive)/p(tests positive)
 = 0.000024825 / 0.0001248 = 0.1989
  

# Wrangling the Billboard Top 100

```{r}
library(readr)
billboard <- read_csv("billboard.csv")
head(billboard,10)
```

## Part A.
```{r}
library(tidyverse)

billboard %>% group_by(performer, song) %>% summarize(count = n()) %>%
  arrange(-count) %>% head(10)
```
This table represents the top 10 most popular songs since 1958. The count is the number of weeks the song appeared on the Billboard top 100 list.

## Part B
```{r}
billboard %>% filter(year > 1958 & year < 2021) %>%
  group_by(year) %>% 
  summarize(num_songs = n_distinct(song)) %>%
  ggplot() + geom_line(aes(year, num_songs))
```

## Part C
```{r}
billboard %>% group_by(song, performer) %>% 
  summarize(max_weeks = max(weeks_on_chart)) %>%
  filter(max_weeks >= 10) %>% 
  group_by(performer) %>% summarize(number_songs_ten_weeks = n()) %>%
  filter(number_songs_ten_weeks >= 30) %>% arrange(-number_songs_ten_weeks) %>%
  ggplot(aes(x = performer, y = number_songs_ten_weeks)) + 
  geom_bar(stat = "identity")
```

  